---
title: 'Reproducible Research: Peer Assessment 1'
author: "Amy Burnett Cross"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---


## Loading and preprocessing the data

Load the data (previously saved in the working directory).
```{r, echo=TRUE}
unzip("activity.zip")
activity <- read.csv("activity.csv")
## Process/transform the data (if necessary) into a format suitable for your analysis
```

Filter all na values.
```{r, echo=TRUE}
activity <- activity[!is.na(activity$steps),]

```


## What is mean total number of steps taken per day?
Calculate the total number of steps taken per day.
```{r, echo=TRUE}
library(dplyr)
daily_activity <- activity %>% group_by(date) %>%
    summarise(total_steps = sum(steps))

head(daily_activity)
```

Print a histogram of the number of steps per day.
```{r, echo=TRUE}
hist(daily_activity$total_steps,
     main = "Histogram of Daily Activity",
     xlab = "Total Steps per Day",
     ylab = "Number of Days")
```

Calculate the mean number of steps per day.
```{r, echo=TRUE}
mean(daily_activity$total_steps)
```

Calculate the median number of steps per day.
```{r, echo=TRUE}
median(daily_activity$total_steps)
```

## What is the average daily activity pattern?

```{r, echo=TRUE}
interval_avg_steps <- activity %>%
    group_by(interval) %>%
    summarise(avg_int_activity = mean(steps))

library(ggplot2)

j <- ggplot(interval_avg_steps, aes(interval, avg_int_activity))
j <- j + geom_line()
j <- j + ggtitle("Average pattern of steps throughout the day")
j <- j + ylab("Average steps")
j

```

Which interval had the maximum average steps?
```{r, echo=TRUE}
max_interval <- interval_avg_steps[which.max(interval_avg_steps$avg_int_activity), ]
print(max_interval)
```

#Imputing missing values


Calculate total number of missing values in dataset.
```{r, echo=TRUE}
activity <- read.csv("activity.csv")

length(which(is.na(activity$steps)))

```

Impute values of missing data with mean steps for the day. Create new data frame that includes imputed values. Now there are no NAs in the data frame.
```{r, echo=TRUE}
activity$steps <- as.numeric(activity$steps)

imputed_activity <- activity %>% 
        group_by(interval) %>%
        mutate(int_avg=mean(steps, na.rm = TRUE),
               imputed_steps = if_else(is.na(steps), int_avg, steps)) %>%
        select(c(imputed_steps, date, interval))
```

Calculate daily steps using imputed values.
```{r, echo=TRUE}
daily_imputed_activity <- imputed_activity %>% group_by(date) %>%
    summarise(total_imputed_steps = sum(imputed_steps))

head(daily_imputed_activity)
```


Create a histogram of the total number of steps taken each day.
```{r, echo=TRUE}
hist(daily_imputed_activity$total_imputed_steps,
     main = "Histogram of Daily Activity with Imputed Values",
     xlab = "Total steps per Day",
     ylab = "Number of Days")
```

Calculate the mean total number of steps taken per day.
```{r, echo=TRUE}
mean(daily_imputed_activity$total_imputed_steps)
```

Calculate the median total number of steps taken per day.
```{r, echo=TRUE}
median(daily_imputed_activity$total_imputed_steps)

```

The mean and median of daily activity differs very little (less than 2 steps per day) when imputed values are included. This makes sense because the method used to impute values mimics daily patterns of activity.


# Differences in activity patterns between weekdays and weekends

```{r, echo=TRUE}
imputed_activity$day_of_week <- weekdays(as.Date(imputed_activity$date))

imputed_activity <- imputed_activity %>% mutate(weekend_status = as.factor(if_else(day_of_week 
                                                                                     %in% (c("Saturday", "Sunday")), 
                                                                                     "Weekend", "Weekday")))



```

Create a panel plot of time series of activity on weekends and weekdays.
```{r, echo=TRUE}
ggplot(imputed_activity, 
       aes(interval, imputed_steps, group = 1)) + 
    geom_line() +
    facet_wrap(~weekend_status, ncol = 1, nrow = 3) +
    ggtitle("Daily Steps") +
    xlab("Interval") +
    scale_x_continuous(breaks = 5) +
    ylab("Total Steps") +
    theme(plot.title = element_text(hjust = 0.5)) +
    theme_classic()
    

# ggplot(IntervalDT , 
    #    aes(x = interval , y = steps, color=`weekday or weekend`)) + 
    # geom_line() + 
    # labs(title = "Avg. Daily Steps by Weektype", x = "Interval", y = "No. of Steps") + 
    # facet_wrap(~`weekday or weekend` , ncol = 1, nrow=2)
```

# ggplot(nmmaps, aes(date,temp))+geom_point(color="aquamarine4")+facet_wrap(~year, nrow=1)





